---
title: "Dataclass_per Units"
author: "David Nora"
date: "2024-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal of the script: This script is part of the paper "The Role of Lithic Technology in Shaping Mobility and Decision-Making: The Case of Ararat-1 Cave". The data present in this script represent the preservation of the artefacts across the sedimentological units at Ararat-1 Cave.

## Load necessary libraries
```{r warning=FALSE}
library(ggplot2)
library(dplyr)
library(readxl)
```

## Load the dataset
```{r warning=FALSE}
file_path <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\raw_data\\Lithic_Attribute_Analysis_Ararat_1_CaveQSR.xlsx"

data <- read_excel(file_path)
```

## Remove rows with NA values
```{r}
data_cleaned <- data %>%
  filter(
    !is.na(DATACLASS),
    !is.na(DEBITAGE),
    !is.na(Merged_UNITs)
  )
```

## Apply sorting order for DEBITAGE and Material Type
```{r}
sorting_order <- c(
  "Core", "Core on flake", "Flake", "Flake<5mm", "Blade", "Bladelet", 
  "Kombewa", "Debordant", "Siret", "Core trimming element", 
  "Burin spall", "Chunk", "Pebble"
)

data_cleaned$DEBITAGE <- factor(data_cleaned$DEBITAGE, levels = sorting_order)
data_cleaned$`Material Type` <- factor(data_cleaned$`Material Type`, levels = c("Obsidian", "Non-Obsidian"))
```

##  Percentages for DATACLASS within each Merged_UNITs
```{r}
dataclass_table <- data_cleaned %>%
  group_by(Merged_UNITs, DATACLASS) %>%
  summarise(Frequency = n(), .groups = "drop") %>%
  group_by(Merged_UNITs) %>%
  mutate(Percentage = (Frequency / sum(Frequency)) * 100) %>%
  arrange(Merged_UNITs, desc(Percentage))

# Calculate percentages for DEBITAGE and DATACLASS combinations within each Merged_UNITs
debitage_dataclass_table <- data_cleaned %>%
  group_by(Merged_UNITs, DEBITAGE, DATACLASS) %>%
  summarise(Frequency = n(), .groups = "drop") %>%
  group_by(Merged_UNITs, DEBITAGE) %>%
  mutate(Percentage = (Frequency / sum(Frequency)) * 100) %>%
  arrange(Merged_UNITs, DEBITAGE, desc(Percentage))

# Display and Save the Tables
print("Percentage Distribution of DATACLASS for Each Merged_UNITs:")
print(dataclass_table)

print("Percentage Distribution of DEBITAGE and DATACLASS for Each Merged_UNITs:")
print(debitage_dataclass_table)

# Save tables as CSV files
write.csv(dataclass_table, "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat_1_cave_lithic_analysis\\Derived data\\dataclass_percentage_table.csv", row.names = FALSE)
write.csv(debitage_dataclass_table, "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat_1_cave_lithic_analysis\\Derived data\\debitage_dataclass_percentage_table.csv", row.names = FALSE)
```

## Generate and Save Plots for Each Merged_UNITs
```{r}
unique_units <- unique(data_cleaned$Merged_UNITs)

# Loop to create and save DATACLASS percentage plots for each Merged_UNITs
for (unit in unique_units) {
  # Filter data for the current unit
  unit_data <- dataclass_table %>% filter(Merged_UNITs == unit)

# Create the plot
  dataclass_plot <- ggplot(unit_data, aes(x = DATACLASS, y = Percentage)) +
    geom_bar(stat = "identity") +
    labs(
      title = paste("DATACLASS Percentages for Merged_UNITs:", unit),
      x = "DATACLASS",
      y = "Percentage (%)"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
    )
  
  # Print the plot
  print(dataclass_plot)
  
  # Save the plot
  ggsave(paste0("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\dataclass_plot_", unit, ".png"), plot = dataclass_plot, width = 10, height = 8, dpi = 300)
}

# Loop to create and save DEBITAGE-DATACLASS percentage plots for each Merged_UNITs
for (unit in unique_units) {
  # Filter data for the current unit
  unit_data <- debitage_dataclass_table %>% filter(Merged_UNITs == unit)
  
  # Create the plot
  debitage_dataclass_plot <- ggplot(unit_data, aes(x = DEBITAGE, y = Percentage, fill = DATACLASS)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    labs(
      title = paste("DEBITAGE and DATACLASS Percentages for Merged_UNITs:", unit),
      x = "DATACLASS",
      y = "Percentage (%)"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14)
    )
  
  # Print the plot
  print(debitage_dataclass_plot)

  # Save the plot
  ggsave(paste0("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\debitage_dataclass_plot_", unit, ".png"), plot = debitage_dataclass_plot, width = 10, height = 8, bg = "white", dpi = 300)
}
```

--- 
End of the Script
