---
title: "Ararat-1 Cave Debitage per Units"
author: "David Nora"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal of the script: This script is part of the paper "The Role of Lithic Technology in Shaping Mobility and Decision-Making: The Case of Ararat-1 Cave". The data present in this script represent the debitage distrubution all sedimentological units at Ararat-1 Cave.

## Load Packages
```{r warning=FALSE}
library(ggplot2)
library(readxl)
library(dplyr)
library(RColorBrewer)
```

# Load the dataset
```{r warning=FALSE}
data_path <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\raw_data\\Lithic_Attribute_Analysis_Ararat_1_CaveQSR.xlsx"
data <- read_excel(data_path)
```

## Define sorting order for debitage types
```{r}
sorting_order <- c("Core", "Core on flake", "Flake", "Flake<5mm", "Blade", "Bladelet", 
                   "Kombewa", "Debordant", "Siret", "Core trimming element", 
                   "Burin spall", "Chunk", "Pebble")

data$DEBITAGE <- factor(data$DEBITAGE, levels = sorting_order)
data$`Merged_UNITs` <- as.factor(data$`Merged_UNITs`)
```


## Aggregate data to count Debitage types within each Merged_UNITs
```{r}
data_summary <- data %>%
  group_by(Merged_UNITs, DEBITAGE) %>%
  summarise(Count = n(), .groups = 'drop')

# View the first few rows of the summarized data
print(head(data_summary))
```

# Generate more colors from the 'Paired' palette
```{r}
palette_extended <- colorRampPalette(brewer.pal(12, "Paired"))

# Assuming you have more than 12 categories, determine the number you need
n_debitage_types <- length(unique(data_summary$DEBITAGE))
extended_colors <- palette_extended(n_debitage_types)
```

# Plotting
```{r}
# Remove rows with NA values in relevant columns
data_clean <- data_summary %>%
  filter(!is.na(DEBITAGE), !is.na(`Merged_UNITs`))

Debitage_Units_Plot <-ggplot(data_clean, aes(x = Merged_UNITs, y = Count, fill = DEBITAGE)) +
  geom_bar(stat = "identity", position = position_dodge(width = 1)) +
  geom_text(aes(label = Count), size = 7, vjust = -0.5, position = position_dodge(width = 1)) +
  labs(title = "Debitage Counts per Merged Units", x = "Merged Units", y = "Count (N)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 18, angle = 0, vjust = 0.5, hjust = 1, face = "bold"),
    axis.text.y = element_text(size = 18, face = "bold"),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    strip.text = element_text(size = 14, face = "bold"),  # Facet labels
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 18, face = "bold"),
  ) +
  scale_fill_manual(values = extended_colors)

# Print the plot to check it
print(Debitage_Units_Plot)
```

## Save the plot
```{r}
ggsave("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\Debitage_Per_Merged_Units.png", plot = Debitage_Units_Plot, width = 25, height = 10, dpi = 300, bg = "white")
```
---

End of the Script
