---
title: "Plots for the number of Dorsal and patterns of Module Flakes"
author: "David Nora"
date: "2025-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal of the Script: The goal of the script is to analyze lithic data from the Ararat 1 Cave by creating visualizations that explore the relationship between different attributes (e.g., scar patterns, number of dorsal scars, and modules of retouch) at Ararat-1 Cave.

```{r}
# Load required libraries
library(readxl)
library(ggplot2)
library(dplyr)
```

## Load the dataset
```{r warning=FALSE}
file_path <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\raw_data\\Lithic_Attribute_Analysis_Ararat_1_CaveQSR.xlsx"

# Read the Excel file for Ararat 1 Cave
ararat_data <- read_excel(file_path)
```

## Filter relevant columns and remove all NA values
```{r}
ararat_filtered <- ararat_data %>%
  filter(
    !is.na(`SCAR PATTERN`) & 
    !is.na(`NO DORSAL SCARS`) & 
    !is.na(`MODULE OF RETOUCH`) &
    !is.na(`Merged_UNITs`) &
    !is.na(`Distance (km)`)
  ) %>%
  select(`SCAR PATTERN`, `NO DORSAL SCARS`, `MODULE OF RETOUCH`, `Merged_UNITs`, `Distance (km)`)

# Remove unused levels from the 'SCAR PATTERN' factor
ararat_filtered$`SCAR PATTERN` <- factor(ararat_filtered$`SCAR PATTERN`)

# Define custom colors for the modules
module_colors <- c("blue", "red", "yellow", "green", "grey")
```

## Create a box plot faceted by Module all Su's
```{r warning=FALSE}
DS_NO_MF <- ggplot(ararat_filtered, aes(x = `SCAR PATTERN`, y = `NO DORSAL SCARS`, fill = `MODULE OF RETOUCH`)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +  # Box plot without outliers
  geom_jitter(aes(color = `SCAR PATTERN`), width = 0.2, size = 1, alpha = 0.7, show.legend = FALSE) +  # Add jittered points
  labs(title = "Box Plot of Number of Dorsal Scars by Scar Pattern all Units",
       x = "Scar Pattern",
       y = "Number of Dorsal Scars",
       fill = "Module of Retouch") +
  scale_fill_manual(values = module_colors) +  # Apply custom module colors
  facet_wrap(~ `MODULE OF RETOUCH`, scales = "fixed", nrow = 1) +  # Use fixed scale across facets
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 12),
    panel.spacing = unit(1, "lines"),
    legend.position = "top"
  ) +
  scale_y_continuous(limits = c(0, 25), breaks = seq(0, 25, by = 5))  # Set consistent y-axis limits and breaks

# Print the plot
print(DS_NO_MF)

# Save the plot
print("Saving plot for all units...")
ggsave(
  filename = "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\DS_MF_Merged_all_Units.png",
  plot = DS_NO_MF,
  width = 10,
  height = 6,
  dpi = 300,
  bg = "white"
)
```

## Function to create the plot for each Merged_UNIT (SU's)
```{r}
create_unit_plot <- function(unit) {
  # Print the current unit being processed
  print(paste("Creating plot for Merged Unit:", unit))
  
  # Filter data for the specific merged unit
  unit_data <- ararat_filtered %>%
    filter(`Merged_UNITs` == unit)  # Use the correct column name
  
  # Create the plot
  plot <- ggplot(unit_data, aes(x = `SCAR PATTERN`, y = `NO DORSAL SCARS`, fill = `MODULE OF RETOUCH`)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7) +  # Box plot without outliers
    geom_jitter(aes(color = `SCAR PATTERN`), width = 0.2, size = 1, alpha = 0.7, show.legend = FALSE) +  # Add jittered points without legend
    labs(title = paste("Box Plot of Number of Dorsal Scars by Scar Pattern - Merged Unit:", unit),
         x = "Scar Pattern",
         y = "Number of Dorsal Scars",
         fill = "Module of Retouch") +
    scale_fill_manual(values = module_colors) +  # Apply custom module colors
    facet_wrap(~ `MODULE OF RETOUCH`, scales = "fixed", nrow = 1) +  # Use fixed scale across facets
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      strip.text = element_text(size = 12),
      panel.spacing = unit(1, "lines"),
      legend.position = "top"
    ) +
    scale_y_continuous(limits = c(0, 25), breaks = seq(0, 25, by = 5))  # Set consistent y-axis limits and breaks
  
  # Print before saving
  print(plot)
  print(paste("Saving plot for Merged Unit:", unit))
  
  # Save the plot
  ggsave(
    filename = paste0("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\DS_MF_Merged_Unit_", unit, ".png"),
    plot = plot, width = 10, height = 6, dpi = 300, bg = 'white'
  )
}
```

## Save the plot for each Merged unit (SU's)
```{r}
# Generate and save plots for each Merged_UNIT
for (unit in unique(ararat_filtered$`Merged_UNITs`)) {
  create_unit_plot(unit)
  print(paste("Plot saved for Merged Unit:", unit))  # Print after saving the plot
}
```

--- 
End of the script
