---
title: "Ararat_1_Cave_MF_RMU_all SU's"
author: "David Nora"
date: "2025-01-24"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal of the Script: The goal of the script is to analyze lithic data from the Ararat 1 Cave by creating visualizations that explore the relationship between different attributes (e.g., Raw material distance, and frequency of modules of retouch) at Ararat-1 Cave.

## Load required libraries
```{r warning=FALSE}
library(readxl)
library(ggplot2)
library(reshape2)
library(dplyr)
```

## Load the dataset
```{r warning=FALSE}
file_path <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\raw_data\\Lithic_Attribute_Analysis_Ararat_1_CaveQSR.xlsx"

# Read the Excel file for Ararat 1 Cave
ararat_data <- read_excel(file_path)
```

## Define the raw material order specific to Ararat 1 Cave in a straight line (Km)
```{r}
raw_material_order_ararat <- c("Gegham", "Geghasar", "Geghasar-ar54", "Geghasar-1", "Geghasar-2", "Hatis",
  "Hatis alpha", "Hatis beta", "Hatis gamma", "Hatis delta", "Gutansar",
  "Tsaghk-ar64", "Tsaghk-2-ar60", "Tsaghk-2 ttavakar", "Sevkar", "Satanakar",
  "Tsaghk-1 kamakar", "Tsaghk-3 damlik", "Pokr arteni", "Pokr arteni-1",
  "Pokr arteni-2", "Mets arteni", "Meydan-dag", "Sarikamis", "Obsidian",
  "Limestone", "Silicified limestone", "Chert", "Chert (brown)", "Chert (green)",
  "Chert (red)", "Calcedony", "Sandstone", "Quartzite", "Metamorphic rock",
  "Basalt", "Mafic lava", "Other")
```

## Clean the dataset
```{r}
# Filter relevant columns and remove rows with missing values
ararat_filtered <- ararat_data[!is.na(ararat_data$Merged_UNITs) & 
                                 !is.na(ararat_data$`MODULE OF RETOUCH`) & 
                                 !is.na(ararat_data$RAWMAT), 
                               c("Merged_UNITs", "MODULE OF RETOUCH", "RAWMAT")]

# Filter relevant columns and remove rows with missing values
ararat_filtered <- ararat_data[!is.na(ararat_data$Merged_UNITs) & 
                                 !is.na(ararat_data$`MODULE OF RETOUCH`) & 
                                 !is.na(ararat_data$RAWMAT), 
                               c("Merged_UNITs", "MODULE OF RETOUCH", "RAWMAT")]

# Function to create faceted bar plot for a specific merged unit
create_faceted_plot <- function(unit, data, raw_material_order) {
  # Filter the data for the specific merged unit
  unit_data <- data[data$Merged_UNITs == unit, ]
  
  # Create a contingency table
  pivot <- as.data.frame(table(unit_data$`MODULE OF RETOUCH`, unit_data$RAWMAT))
  names(pivot) <- c("Module", "RawMaterial", "Count")
  
  # Filter only for existing raw materials in the defined order
  pivot <- pivot[pivot$RawMaterial %in% raw_material_order, ]
  
  # Calculate percentages within each Module
  pivot <- pivot %>%
    group_by(Module) %>%
    mutate(Percentage = Count / sum(Count) * 100) %>%
    ungroup()
  
  # Remove rows with 0 counts to avoid 0% labels
  pivot <- pivot %>%
    filter(Count > 0)
  
  # Set the order of the raw materials
  pivot$RawMaterial <- factor(pivot$RawMaterial, levels = raw_material_order)
  
  # Convert "Module" to a factor for faceting and coloring
  pivot$Module <- factor(pivot$Module)
  
  # Create the faceted bar plot

  ggplot(pivot, aes(x = Count, y = RawMaterial, fill = Module)) +
    geom_bar(stat = "identity", position = "stack") +
    geom_text(aes(label = paste0(Count, " (", round(Percentage, 1), "%)")),
              position = position_stack(vjust = 0.5), hjust = -0.1, size = 5) +  # Add labels
    labs(title = paste("Module Flakes per Raw Material - Merged Unit", unit),
         x = "Count (N)", y = "Raw Material", fill = "Module") +
    theme_minimal() +
    scale_fill_manual(values = c("blue", "red", "gold", "green", "grey")) +  # Custom colors for Modules
    scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, by = 5)) +  # Fix X-axis
    facet_wrap(~ Module, scales = "free_x", nrow = 1) +  # Facet by Module
    theme(
      strip.text = element_text(size = 14),  # Facet labels font size
      panel.spacing = unit(1, "lines"),
      legend.text  = element_text(size = 14),
      legend.title = element_text(size = 16),
      axis.title.x = element_text(size = 12, face = "bold"),
      axis.text.x  = element_text(angle = 0, hjust = 1, size = 10, face = "bold"),
      axis.title.y = element_text(size = 14, face = "bold"),
      axis.text.y  = element_text(size = 12, face = "bold") # Increase spacing between facets
    )
}
```

## Loop through all unique merged units, generate plots, and save
```{r}
output_directory <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data"
unique_units <- unique(ararat_filtered$Merged_UNITs)
unit_plots <- list()

for (unit in unique_units) {
  # Create the plot
  plot <- create_faceted_plot(unit, ararat_filtered, raw_material_order_ararat)
  
  # Save the plot as a PNG file
  file_name <- paste0(output_directory, "\\Module_Flakes_Unit_", unit, ".png")
  ggsave(filename = file_name, plot = plot, width = 14, height = 6, bg = 'white', dpi = 300)
  
  # Store the plot in a list
  unit_plots[[paste0("Unit_", unit)]] <- plot
}

# Optionally, display all plots
for (plot_name in names(unit_plots)) {
  print(unit_plots[[plot_name]])
}
```

---
End of the script
