---
title: "Module Flakes stats"
author: "David Nora"
date: "2025-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal of the script:This script is part of the paper "The Role of Lithic Technology in Shaping Mobility and Decision-Making: The Case of Ararat-1 Cave". The data present in this script represent the metrics technological length and width and thickness on module flakes categories for each sedimentological units at Ararat-1 Cave.

## Load Packages
```{r warning=FALSE}
library(R.utils)
library(ggplot2)
library(tidyverse)
library(ggpubr)
```

## Load Dataset and Remove NA
```{r warning=FALSE}
Ararat_1 <- readxl::read_excel("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\raw_data\\Lithic_Attribute_Analysis_Ararat_1_CaveQSR.xlsx") %>%
  filter(
    !is.na(`MODULE OF RETOUCH`) & 
    !is.na(`TECHNOLOGICAL LENGTH`) & 
    !is.na(`TECHNOLOGICAL WIDTH`) & 
    !is.na(THICKNESS) & 
    !is.na(`Merged_UNITs`)
  )
```

## Define Custom Colors for Modules
```{r}
module_colors <- c("blue", "red", "gold", "green", "grey")
names(module_colors) <- c("Module 0", "Module 1", "Module 2", "Module 3", "Module 4")
```

## Summary Statistics Function by Module
```{r}
compute_summary_stats_FM <- function(data) {
  data %>%
    group_by(`MODULE OF RETOUCH`) %>%
    summarise(
      Mean_Length = mean(`TECHNOLOGICAL LENGTH`, na.rm = TRUE),
      SD_Length = sd(`TECHNOLOGICAL LENGTH`, na.rm = TRUE),
      Median_Length = median(`TECHNOLOGICAL LENGTH`, na.rm = TRUE),
      Min_Length = min(`TECHNOLOGICAL LENGTH`, na.rm = TRUE),
      Max_Length = max(`TECHNOLOGICAL LENGTH`, na.rm = TRUE),
      
      Mean_Width = mean(`TECHNOLOGICAL WIDTH`, na.rm = TRUE),
      SD_Width = sd(`TECHNOLOGICAL WIDTH`, na.rm = TRUE),
      Median_Width = median(`TECHNOLOGICAL WIDTH`, na.rm = TRUE),
      Min_Width = min(`TECHNOLOGICAL WIDTH`, na.rm = TRUE),
      Max_Width = max(`TECHNOLOGICAL WIDTH`, na.rm = TRUE),
      
      Mean_Thickness = mean(THICKNESS, na.rm = TRUE),
      SD_Thickness = sd(THICKNESS, na.rm = TRUE),
      Median_Thickness = median(THICKNESS, na.rm = TRUE),
      Min_Thickness = min(THICKNESS, na.rm = TRUE),
      Max_Thickness = max(THICKNESS, na.rm = TRUE)
    )
}
```

## Run the Function for the Dataset
```{r}
summary_stats_FM_Ara_1 <- compute_summary_stats_FM(Ararat_1)

# Print the Summary
print(summary_stats_FM_Ara_1)
```

## Reshape Dataset into Long Format
```{r}
reshape_data <- function(data) {
  data %>%
    pivot_longer(
      cols = c(`TECHNOLOGICAL LENGTH`, `TECHNOLOGICAL WIDTH`, `THICKNESS`),
      names_to = "Variable",
      values_to = "Value"
    ) %>%
    mutate(Variable = recode(Variable,
                             "TECHNOLOGICAL LENGTH" = "TL",
                             "TECHNOLOGICAL WIDTH"  = "TW",
                             "THICKNESS"            = "TH"))
}

Ararat_1_Cave_long <- reshape_data(Ararat_1)
```

## Create the function for each SU and combined SU's
```{r}
# Function to create a faceted boxplot for a single unit
create_faceted_boxplot_per_unit <- function(data, unit, dataset_name) {
  ggplot(data %>% filter(`Merged_UNITs` == unit), 
         aes(x = Variable, y = Value, color = `MODULE OF RETOUCH`)) +
    geom_jitter(aes(group = interaction(Variable, `MODULE OF RETOUCH`)), 
                width = 0.15, size = 1, alpha = 0.5, show.legend = FALSE) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7, fill = "white") +
    labs(
      title = paste("Boxplots of Metrics by Module - Unit:", unit, "-", dataset_name),
      x = "Metric",
      y = "Value (mm)"
    ) +
    ylim(0, 30) +
    facet_wrap(~ `MODULE OF RETOUCH`, nrow = 1) +  # Arrange facets in one row (side by side)
    theme_minimal() +
    theme(
      legend.text  = element_text(size = 14),
      legend.title = element_text(size = 16),
      axis.title.x = element_text(size = 14, face = "bold"),
      axis.text.x  = element_text(size = 10, face = "bold"),
      axis.title.y = element_text(size = 14, face = "bold"),
      axis.text.y  = element_text(size = 14, face = "bold")
    ) +
    scale_color_manual(values = module_colors)
}

# Function to create a faceted boxplot using all data (no unit filtering)
create_faceted_boxplot_all_data <- function(data, dataset_name) {
  ggplot(data, aes(x = Variable, y = Value, color = `MODULE OF RETOUCH`)) +
    geom_jitter(aes(group = interaction(Variable, `MODULE OF RETOUCH`)), 
                width = 0.15, size = 1, alpha = 0.5, show.legend = FALSE) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7, fill = "white") +
    labs(
      title = paste("Boxplots of Metrics by Module - All Data Combined -", dataset_name),
      x = "Metric",
      y = "Value (mm)"
    ) +
    ylim(0, 30) +
    facet_wrap(~ `MODULE OF RETOUCH`, nrow = 1) +  # Arrange facets in one row (side by side)
    theme_minimal() +
    theme(
      legend.text  = element_text(size = 14),
      legend.title = element_text(size = 16),
      axis.title.x = element_text(size = 14, face = "bold"),
      axis.text.x  = element_text(size = 10, face = "bold"),
      axis.title.y = element_text(size = 14, face = "bold"),
      axis.text.y  = element_text(size = 14, face = "bold")
    ) +
    scale_color_manual(values = module_colors)
}
```

## Plot and Save
```{r}
for (unit in unique(Ararat_1_Cave_long$`Merged_UNITs`)) {
  plot <- create_faceted_boxplot_per_unit(Ararat_1_Cave_long, unit, "Ararat_1_Cave")
  print(plot)
  ggsave(
    filename = paste0(
      "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\Combined_Faceted_Boxplot_By_Unit_", 
      unit, "_Ararat_1_Cave.png"
    ),
    plot = plot, width = 12, height = 8, dpi = 300, bg = "white"
  )
}

combined_all_data_plot <- create_faceted_boxplot_all_data(Ararat_1_Cave_long, "Ararat_1_Cave")
print(combined_all_data_plot)
ggsave(
  filename = "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\Combined_Faceted_Boxplot_All_Data_Ararat_1_Cave.png",
  plot = combined_all_data_plot, width = 12, height = 8, dpi = 300, bg = "white"
)

```
---
End of the script
