---
title: "Retouched pieces per SU"
author: "David Nora"
date: "2025-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal of the Script: The goal of the script is to analyze lithic data from the Ararat 1 Cave by creating visualizations that explore the relationship between retouched pieces, e.g., Raw material distance, and frequency of retouch pieces at Ararat-1 Cave.

## Load Libraries
```{r warning=FALSE}
# Load required libraries
library(readxl)
library(ggplot2)
library(dplyr)
```

## Load the dataset
```{r warning=FALSE}
file_path <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat_1_cave_lithic_analysis\\Raw data\\Lithic_Attribute_Analysis_Ararat_1_CaveQSR.xlsx"
ararat_data <- read_excel(file_path)
```

## Define the raw material order specific to Ararat 1 Cave in a straight line (Km)
```{r}
raw_material_order_ararat <- c("Gegham", "Geghasar", "Geghasar-ar54", "Geghasar-1", "Geghasar-2", "Hatis",
  "Hatis alpha", "Hatis beta", "Hatis gamma", "Hatis delta", "Gutansar",
  "Tsaghk-ar64", "Tsaghk-2-ar60", "Tsaghk-2 ttavakar", "Sevkar", "Satanakar",
  "Tsaghk-1 kamakar", "Tsaghk-3 damlik", "Pokr arteni", "Pokr arteni-1",
  "Pokr arteni-2", "Mets arteni", "Meydan-dag", "Sarikamis", "Obsidian",
  "Limestone", "Silicified limestone", "Chert", "Chert (brown)", "Chert (green)",
  "Chert (red)", "Calcedony", "Sandstone", "Quartzite", "Metamorphic rock",
  "Basalt", "Mafic lava", "Other")
```

# Filter relevant columns and remove rows with missing values
```{r}
ararat_data <- ararat_data %>%
  filter(!`TYPOLOGY A` %in% c("38. Naturally-backed knife", 
                              "1. Typical Levallois flake", 
                              "5. Pseudo-Levallois point"))

ararat_filtered <- ararat_data[!is.na(ararat_data$Merged_UNITs) & 
                               !is.na(ararat_data$`TYPOLOGY A`) & 
                               !is.na(ararat_data$RAWMAT), 
                               c("Merged_UNITs", "TYPOLOGY A", "RAWMAT")]
```

## Create a function for bar plots
```{r}
create_typology_plot <- function(unit, data, raw_material_order) {
  # Filter data for the specific merged unit
  unit_data <- data[data$Merged_UNITs == unit, ]
  
  # Create a contingency table
  pivot <- as.data.frame(table(unit_data$`TYPOLOGY A`, unit_data$RAWMAT))
  names(pivot) <- c("Typology", "RawMaterial", "Count")
  
  # Filter raw materials based on the defined order
  pivot <- pivot[pivot$RawMaterial %in% raw_material_order, ]
  
  # Remove rows with zero counts
  pivot <- pivot[pivot$Count > 0, ]
  
  # Set raw material order
  pivot$RawMaterial <- factor(pivot$RawMaterial, levels = raw_material_order)
  
  # Plot the bar chart
  ggplot(pivot, aes(x = Count, y = RawMaterial, fill = Typology)) +
    geom_bar(stat = "identity", position = "stack") +
    geom_text(aes(label = Count), position = position_stack(vjust = 0.5), size = 3) +
    labs(
      title = paste("Raw Material Counts by Typology - Merged Unit", unit),
      x = "Count", y = "Raw Material", fill = "Typology"
    ) +
    theme_minimal() +
    scale_x_continuous(limits = c(0, 5), breaks = 0:5) +
      theme(
        axis.text.x   = element_text(angle = 0, hjust = 1, size = 14, face = "bold"),
        axis.title.x  = element_text(size = 16),
        axis.text.y   = element_text(size = 14, face = "bold"),
        axis.title.y  = element_text(size = 16),
        legend.text   = element_text(size = 16),
        legend.title = element_text(size = 18),
      strip.text = element_text(size = 12),
      panel.spacing = unit(1, "lines"),
      color = "Typology"
    )
}
```

```{r}
# Define output directory and ensure it exists
output_directory <- "C:/Users/David/OneDrive - huji.ac.il/Desktop/Archaeology/Publications/In_Proguess/PhD/QSR/Ararat_1_cave_lithic_analysis/Derived data"
if (!dir.exists(output_directory)) {
  dir.create(output_directory, recursive = TRUE)
}

# Loop through unique units and generate/save plots
# Convert Merged_UNITs to character (in case they are factors)
unique_units <- unique(as.character(ararat_filtered$Merged_UNITs))
typology_plots <- list()

for (unit in unique_units) {
  # Create the plot
  plot_obj <- create_typology_plot(unit, ararat_filtered, raw_material_order_ararat)
  
  # Save the plot as a PNG file
  file_name <- paste0(output_directory, "/Typology_Unit_", unit, ".png")
  ggsave(filename = file_name, plot = plot_obj, width = 12, height = 6, bg = "white", dpi = 300)
  
  # Store the plot in a list
  typology_plots[[unit]] <- plot_obj
}

# (Optional) Print a message with the saved file names
saved_files <- sapply(unique_units, function(unit) paste0("Typology_Unit_", unit, ".png"))
cat("Saved files:\n", paste(saved_files, collapse = "\n"))
```

```{r}
lapply(typology_plots,print)
```

---
End of script 
