---
title: "Ararat-1 Cave Raw Materials per Unit"
author: "David Nora"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal of the script: This script is part of the paper "The Role of Lithic Technology in Shaping Mobility and Decision-Making: The Case of Ararat-1 Cave". The data present in this script represent the raw materials distrubution across the sedimentological units of Ararat-1 Cave.

## Load Packages
```{r warning=FALSE}
library(ggplot2)
library(readxl)
library(dplyr)
```

## Load the dataset
```{r warning=FALSE}
file_path <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\raw_data\\Lithic_Attribute_Analysis_Ararat_1_CaveQSR.xlsx"

data <- read_excel(file_path)
```

## Define sorting order for RAWMAT
```{r}
sorting_order <- c(
  "Gegham", "Geghasar", "Geghasar-ar54", "Geghasar-1", "Geghasar-2", "Hatis",
  "Hatis alpha", "Hatis beta", "Hatis gamma", "Hatis delta", "Gutansar",
  "Tsaghk-ar64", "Tsaghk-2-ar60", "Tsaghk-2 ttavakar", "Sevkar", "Satanakar",
  "Tsaghk-1 kamakar", "Tsaghk-3 damlik", "Pokr arteni", "Pokr arteni-1",
  "Pokr arteni-2", "Mets arteni", "Meydan-dag", "Sarikamis", "Obsidian",
  "Limestone", "Silicified limestone", "Chert", "Chert (brown)", "Chert (green)",
  "Chert (red)", "Calcedony", "Sandstone", "Quartzite", "Metamorphic rock",
  "Basalt", "Mafic lava", "Other"
)
```

## Convert factors
```{r}
data$RAWMAT <- factor(data$RAWMAT, levels = sorting_order)
data$`Merged_UNITs` <- as.factor(data$`Merged_UNITs`)
data$`Material Type` <- factor(data$`Material Type`, levels = c("Obsidian", "Non-Obsidian"))
```

##Remove NA
```{r}
data_clean <- data %>%
  filter(!is.na(RAWMAT), !is.na(`Material Type`), !is.na(`Merged_UNITs`))
```

## Data Visualization
```{r warning=FALSE}
# Create a custom labeller function
custom_labeller <- function(unit) {
  paste("S.U", unit)
}

# Apply the custom labeller in facet_wrap
p <- ggplot(data_clean, aes(x = RAWMAT, y = ..count.., fill = `Material Type`)) +
  geom_bar(stat = "count", position = "dodge", width = 0.6) +
  geom_text(
    stat = "count", 
    aes(label = ..count..), 
    vjust = -0.5, 
    position = position_dodge(width = 0.6), 
    size = 3
  ) +
  labs(
    title = "Distribution of Raw Materials per Unit",
    x = "Raw Material",
    y = "Count (N)",
    fill = "Material Type"
  ) +
  facet_wrap(~`Merged_UNITs`, scales = "free_y", ncol = 1, labeller = labeller(`Merged_UNITs` = custom_labeller)) +  # Add custom labels
  scale_y_continuous(expand = expansion(mult = c(0, 0.3))) +  # Add space above bars
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, angle = 90, vjust = 0.5, hjust = 1),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    strip.text = element_text(size = 12, face = "bold"),  # Facet labels
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold")
  )

# Print the plot
print(p)

```

## Save the Plot
```{r}
ggsave("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\P_RMU_Units.png", plot = p, width = 12, height = 8, dpi = 300, bg = 'white')
```

----
End of Script
