---
title: "Debitage_TL_vs_TW_all_units"
author: "David Nora"
date: "2025-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal of the script: This script is part of the paper "The Role of Lithic Technology in Shaping Mobility and Decision-Making: The Case of Ararat-1 Cave". The data present in this script represent the metrics technological length and width of all sedimentological units in one plot filled by the debitage categories, and divied between obsidian and non-obsidian at Ararat-1 Cave.

## Load Required Libraries
```{r}
library(ggplot2)
library(dplyr)
library(readxl)
```

## Load the dataset
```{r warning=FALSE}
data_path <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\raw_data\\Lithic_Attribute_Analysis_Ararat_1_CaveQSR.xlsx"
Ararat_1_Cave <- read_excel(data_path)
```

## Prepare the data
```{r}
Ararat_1_Cave <- Ararat_1_Cave %>%
  filter(!is.na(DEBITAGE))
```

## Define sorting order for debitage types
```{r}
sorting_order <- c("Core", "Core on flake", "Flake", "Flake<5mm", "Blade", "Bladelet", 
                   "Kombewa", "Debordant", "Siret", "Core trimming element", 
                   "Burin spall", "Chunk", "Pebble")

Ararat_1_Cave$DEBITAGE <- factor(Ararat_1_Cave$DEBITAGE, levels = sorting_order)
Ararat_1_Cave$`Material Type` <- factor(Ararat_1_Cave$`Material Type`, levels = c("Obsidian", "Non-Obsidian"))
```

## Define the plot for all units combined
```{r warning=FALSE}
P <-ggplot(Ararat_1_Cave, aes(x = `TECHNOLOGICAL WIDTH`, y = `TECHNOLOGICAL LENGTH`, fill = `DEBITAGE`)) +
  geom_point(shape = 21, color = "black", alpha = 0.6, size = 3) +
  scale_fill_manual(values = c("#e6194B", "#3cb44b", "#ffe119", "#4363d8", "#f58231", 
                               "#911eb4", "#46f0f0", "#f032e6", "#bcf60c", "#fabebe",
                               "#008080", "#e6beff", "#9A6324")) + # Custom colors
  labs(title = "Technological Length vs. Width for All Debitage Categories",
       x = "Technological Width (mm)", 
       y = "Technological Length (mm)", 
       fill = "Debitage Category") +
  scale_x_continuous(limits = c(0, 80)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12, angle = 0, hjust = 1),
    axis.text.y = element_text(size = 12, angle = 0, hjust = 1),
    legend.position = "right",
    legend.text = element_text(face =  "bold", size = 14),
    legend.title = element_text(size = 14),
    legend.key.size = unit(1.5, "lines"),
    strip.text = element_text(face = "bold", size = 14)
  ) +
  facet_wrap(~ `Material Type`, ncol = 2) # Split into two plots for Obsidian and Non-Obsidian
print(P)
```

## Save the Plot
```{r warning=FALSE}
ggsave("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\plot_debita_Obsidian_nonObsidian.png0_debitage_.png", plot = P, width = 14, height = 8, dpi = 300, bg = 'white')
```

---
End of the script
