---
title: "Retouched Pieces width/length Obsidian vs. Non-Obsidian"
author: "David Nora"
date: "2024-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal of the script: This script is part of the paper "The Role of Lithic Technology in Shaping Mobility and Decision-Making: The Case of Ararat-1 Cave". The data present in this script represent the metrics technological length and width of the sedimentological units 3 and 4 by Obsidian and non-obsidian, as well the presence of retouched pieces at Ararat-1 Cave.

## Load Required Libraries
```{r warning=FALSE}
library(ggplot2)
library(dplyr)
library(readxl)
```

## Load the dataset
```{r warning=FALSE}
data_path <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\raw_data\\Lithic_Attribute_Analysis_Ararat_1_CaveQSR.xlsx"
Ararat_1_Cave <- read_excel(data_path)
```

## Prepare the data
```{r}
#remove non retouched typologies
Ararat_1_Cave <- Ararat_1_Cave %>%
  filter(!`TYPOLOGY A` %in% c("38. Naturally-backed knife", 
                              "1. Typical Levallois flake", 
                              "5. Pseudo-Levallois point"))

# Ensure numerical conversion if they are not already and handle NAs
Ararat_1_Cave$`TECHNOLOGICAL LENGTH` <- as.numeric(Ararat_1_Cave$`TECHNOLOGICAL LENGTH`)
Ararat_1_Cave$`TECHNOLOGICAL WIDTH` <- as.numeric(Ararat_1_Cave$`TECHNOLOGICAL WIDTH`)

# Create a column to indicate whether the data point is Typology A, handle NAs
Ararat_1_Cave$Is_Typology_A <- ifelse(Ararat_1_Cave$`TYPOLOGY A` == "A", "Typology A", "Not Typology A")
Ararat_1_Cave$Is_Typology_A[is.na(Ararat_1_Cave$`TYPOLOGY A`)] <- "NA"

# Filter out rows with NA in key plotting columns if necessary
Ararat_1_Cave <- Ararat_1_Cave %>%
  filter(!is.na(`TECHNOLOGICAL LENGTH`) & !is.na(`TECHNOLOGICAL WIDTH`))
```

## Define the plot function
```{r}
plots <- list()
for (unit in unique(Ararat_1_Cave$`Merged_UNITs`)) {
  unit_data <- filter(Ararat_1_Cave, `Merged_UNITs` == unit)
  plot_title <- paste("Unit", unit, ": Technological Length vs Width")
  plots[[as.character(unit)]] <- ggplot(unit_data, aes(x = `TECHNOLOGICAL WIDTH`, y = `TECHNOLOGICAL LENGTH`, color = `Material Type`, shape = Is_Typology_A)) +
    geom_point(alpha = 0.6, size = 4) +
    scale_color_manual(values = c("Obsidian" = "red", "Non-Obsidian" = "blue")) +
    scale_shape_manual(values = c("Typology A" = 17, "Not Typology A" = 16, "NA" = 21),
                       labels = c("Typology A" = "Typology A", "Not Typology A" = "Yes", "NA" = "No")) +
    labs(title = plot_title, x = "Technological Width (mm)", y = "Technological Length (mm)", shape = "Retouched Pieces") +
  scale_x_continuous(limits = c(0, 80)) +
   theme_minimal() +
  theme(
    axis.text.x    = element_text(angle = 0, hjust = 1),
    legend.position = "right",
    legend.text     = element_text(size = 16),
    legend.title    = element_text(size = 18),
    legend.key.size = unit(1.5, "lines"))
}
```

## Print plots for each unit
```{r}
for (unit in names(plots)) {
  print(plots[[unit]])
}
```

##  Save the plots
```{r}
# Save plots for each unit
output_dir <- "C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\"
for (unit in names(plots)) {
  plot_path <- paste0(output_dir, "Unit_", unit, "_Technological_Length_vs_Width.png")
  ggsave(filename = plot_path, plot = plots[[unit]], width = 8, height = 6, dpi = 300, bg = "white")
}

```
---
End of Script
