---
title: "Vertical Distribution of the lithic assemblage at Ararat-1 Cave"
author: "David Nora"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal of script: This script aims to show the vertical distribution of the lithic assemblage in a scatterplot visualization. In the same plot, we overstack the length and width of the morphological metric to understand if there is any first evidence of sorting throughout the stratigraphy. we also present an approximate size for each lithic and display it according to the bottom elevation (the contact of the sedimentary units).

## Load necessary libraries
```{r warning=FALSE}
library(ggplot2)
library(dplyr)
library(readxl)
library(cowplot)
```


## Load and tidy the dataset
```{r warning=FALSE}
# Load the dataset
data_path <- note

Ararat_1_Cave <- read_excel(data_path)

# Ensure necessary columns are numeric; replace non-numeric values with NA
Ararat_1_Cave <- Ararat_1_Cave %>%
  mutate(
    BOTTOMELEVATION = as.numeric(BOTTOMELEVATION),
    `MORPHOLOGICAL LENGTH` = as.numeric(`MORPHOLOGICAL LENGTH`),
    `MORPHOLOGICAL WIDTH` = as.numeric(`MORPHOLOGICAL WIDTH`)
  )

# Clean and standardize Material.Type
Ararat_1_Cave$`Material Type` <- Ararat_1_Cave$`Material Type` %>%
  trimws() %>%               # Remove leading and trailing spaces
  tolower() %>%              # Convert to lowercase
  tools::toTitleCase()       # Capitalize first letters

# Filter and process data
filtered_Ararat_1_Cave <- Ararat_1_Cave %>%
  filter(
    !is.na(BOTTOMELEVATION) &
    !is.na(`MORPHOLOGICAL LENGTH`) &
    !is.na(`MORPHOLOGICAL WIDTH`) &
    `MORPHOLOGICAL LENGTH` > 0 &
    `MORPHOLOGICAL WIDTH` > 0
  ) %>%
  mutate(
    `Approximate Size` = `MORPHOLOGICAL LENGTH` * `MORPHOLOGICAL WIDTH`,
    `Material Type` = ifelse(`Material Type` == "Obsidian", "Obsidian", "Non-Obsidian")
  )
```

## Plot 1: Bottom Elevation vs Morphological Length and Width
```{r warning=FALSE}
# Create separate datasets for Length and Width
length_data <- filtered_Ararat_1_Cave %>%
  select(BOTTOMELEVATION, `MORPHOLOGICAL LENGTH`, `Material Type`) %>%
  rename(Metric = `MORPHOLOGICAL LENGTH`) %>%
  mutate(Measurement = "Length")

width_data <- filtered_Ararat_1_Cave %>%
  select(BOTTOMELEVATION, `MORPHOLOGICAL WIDTH`, `Material Type`) %>%
  rename(Metric = `MORPHOLOGICAL WIDTH`) %>%
  mutate(Measurement = "Width")

# Combine length and width datasets for plotting
combined_data <- bind_rows(length_data, width_data)

# Plot
Plot1 <- ggplot(combined_data, aes(x = Metric, y = BOTTOMELEVATION, color = `Material Type`, shape = Measurement)) +
  geom_point(alpha = 0.8) +
  scale_y_continuous(limits = c(97.80, 100.10), breaks = seq(97.80, 100.10, by = 0.10)) +
  labs(
    title = "Bottom Elevation by Morphological Length and Width (10 cm Intervals)",
    x = "Morphological Measurements (Length and Width) in mm",
    y = "Bottom Elevation (m)",
    color = "Material Type",
    shape = "Measurement Type"
  ) +
  theme_minimal() +
  theme(
  axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16, face = "bold"),
)
print(Plot1)
```

## Plot 2: Bottom Elevation vs Approximate Size
```{r warning=FALSE}
# Plot Approximate Size
Plot2 <- ggplot(filtered_Ararat_1_Cave, aes(x = `Approximate Size`, y = BOTTOMELEVATION, color = `Material Type`)) +
  geom_point(alpha = 0.7) +
  scale_y_continuous(limits = c(97.80, 100.10), breaks = seq(97.80, 100.10, by = 0.10)) +
  labs(
    title = "Bottom Elevation by Approximate Area (Separated by Material)",
    x = "Approximate Size (Length x Width) in mm²",
    y = "Bottom Elevation (m)",
    color = "Material Type"
  ) +
  theme_minimal() +
  theme(
  axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16, face = "bold"),
)
print(Plot2)
```

## Save Plots
```{r}
# Save the plots
ggsave("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\Plot1_Bottom_Elevation)L_W.png", plot = Plot1, width = 8, height = 10, dpi = 300, bg = 'white')

ggsave("C:\\Users\\David\\OneDrive - huji.ac.il\\Desktop\\Archaeology\\Publications\\In_Proguess\\PhD\\QSR\\Ararat1caveLithicQSR\\analysis\\data\\derived_data\\Plot2_Bottom_Elevation)L_W.png", plot = Plot2, width = 8, height = 10, dpi = 300, bg = 'white')
```
